---
title: "Dmel Social Environment on Behaviour"
author: "Erin Macartney"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
      code_folding: show
      toc_depth: 3
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = F}
library(tidyverse)
library(lme4)
library(lmerTest)
library(here) # write out the path
library(kableExtra)
library(sjlabelled)
library(ggpubr)
library(rstatix)
library(car)
library(RColorBrewer)
library(readxl)

# col_accent1 = "slateblue1"
# col_accent2 = "springgreen3"
# col_detail1 = "tomato2"
```

# Locomotion

## Data parsing

First let's test-load one data file to see how to trim it into relevant bits. Definition in this files is so that first 4 rows and first 6 columns are redundant.

The first portion loads and parses the raw data file lines.

```{r 1_data_load}
data_path <- here("Data/locomotion/")
data_files <- list.files(here("Data/locomotion/"), recursive = T)
data_files[1:24]

#Loading in session data
Session_data <- read_excel(here("Data/session_data/Session_data_extended.xlsx"))
```

Currently each file represents a set of uniquely analyzed individuals and hence no additional complications arise (like, e.g., having two files with the same individual assayed on two occasions).

In the next step we test the approach on several steps. First - we process a sample CSV file, which is unstructured and contains a lot of spurious lines and data (e.g., control variables and comments generated by the unit).

```{r 2_parse_lines}
# read in one specific file for Mac users 
#dat_temp <- readLines(paste(data_path, data_files[1], sep = ''))

# # read in one specific file for PC users 
dat_temp <- readLines(paste(data_path, data_files[1], sep = '/'))

glimpse(dat_temp)

```

The loaded data is just a vector of strings, each being a line from the original CSV file. In the next chunk we skip the first 4 lines (fixed number, lines containing technical parameters of the units), and load 5 lines (which excludes the header line - so in fact 6 lines), skipping the last technical line. At the end we clean the file (removing empty columns `2:6` and wells `c(F6, F7, F8)` - they are always empty in our system). Adjust as needed. The resulting file contains each assayed well as separate column, and for each there are 5 times intervals of locomotion monitoring.

```{r 3_read_metadata}
dat_temp_df <- read.csv(text = dat_temp, header = T, sep = ',',
                        quote = '\"', dec = '.', skip = 4, nrows = length(dat_temp) - 4 - 2,
                        stringsAsFactors = F)

dat_temp_df <- dat_temp_df[, -(2:6)] # remove spurious columns
dat_temp_df <- dat_temp_df %>% select(!(F3:F8)) #remove wells that do not contain data
dat_temp_df

glimpse(dat_temp_df)
```

Here we extract and append the run (subject) ID.

```{r 4_parse_runID}
head_temp <- readLines(paste(data_path, data_files[1], sep = '/'), n = 4)

# using RE - this method is more flexible as the structure and location of ID can change

# both of the below definitions will work, but second is more precise
# it uses the exact format of the run ID

id_index <- grep('Subject Identification', head_temp)
# id_index <- grep('.*Subject Identification\\\",\"([A-Z0-9]{7}_B_[0-9]{6}_[0-9]{1,2})\\\"', head_temp)
run_id <- gsub('.*Subject Identification\\\",\"([A-Z]{1}[0-9]{3})\\\"', '\\1', head_temp[id_index])
dat_temp_df$Datafile_ID <- run_id

# assay_date <- str_sub(run_id, 11, 16)
# dat_temp_df$date <- as.Date(assay_date, format = "%y%m%d")
glimpse(dat_temp_df)
```
Change to long format. Now each well x time combination has it's separate row, and since we have datafile ID we can link wells to specific individuals - e.g., to link-in sex information. We also rename variable according to convention: categorical variables - `Names_start_with_upper_case`; continuous variables - `all_lower_case`.

```{r 5_turn_to_long}
dat_temp_df2 <-
  dat_temp_df %>%
  pivot_longer(names_to = 'well_id', values_to = 'arena_distance', cols = matches('[A-H][1-9]')) %>%
  rename(time = TIME, temperature = TEMPERATURE, round = ROUND,
         variable = VARIABLE)

# head(dat_temp_df2)
# glimpse(dat_temp_df2)
```

We have to add the `Individuals_ID` variable to be able to link locomotion data to sex data. It will serve to link Datafile_ID with respective Individuals_ID - and through it with appropriate well number. The run register contains several variables used to group individuals into several blocks that may share some of the systematic variation.

### Sample data frame

### Locomotion

```{r 7_merge_metadata}
# run_register <- read_delim(here('Data', 'run_data', 'run_register_pilot2.csv'), delim = ';')
# glimpse(run_register)

#need to add an column called Individual_ID that combined assay name and well number to be able to link session data to dataframe

#Session_data
Session_data$Individual_ID <- paste(Session_data$Assay_ID,Session_data$Position, sep = "_")

#data_temp
dat_temp_df2$Individual_ID <- paste(dat_temp_df2$Datafile_ID, dat_temp_df2$well_id, sep = "_")


dat_temp_df3 <- dat_temp_df2 %>%
  left_join(Session_data, by = "Individual_ID")

```

### Ymaze

### Habituation

